---
title: "CodeBook"
author: "Angel Mancera"
date: "6/1/2020"
output: pdf_document
---

## Course Project
Hereby we will understand the variables of the final data set. 

## Initial Files
After downloading and unzipping the zip file we can recognize six files that will be useful: 
x_train.txt, x_test.txt, y_train.txt, y_test.txt, subject_train.txt and subject_test.txt. 

These files contain the data and information for naming the variables correctly. 
We use specifically the features file to label correctly both x_something files 
as it contains the good variable name. The same is for the activity file with the 
y_something files. 

## The Inner-Workings of the Project
Firstly, we download the files using the download.file function, and the two if statements:
```{r setup}
files <- "getdata_projectfiles_UCI HAR Dataset.zip"

if (!file.exists(files)){
    fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
    download.file(fileURL, files, method="curl")
}  

if (!file.exists("UCI HAR Dataset")) { 
    unzip(files) 
}
```
These essentially check if the file has already been downloaded, and if not, it downloads it!

Then, we read every data table: 
``` {r follow-up}
feat <- read.table("UCI HAR Dataset/features.txt", col.names = c("n","functions"))
actvs <- read.table("UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject.ta <- read.table("UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
subject.te <- read.table("UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
X.te <- read.table("UCI HAR Dataset/test/X_test.txt", col.names = feat$functions)
y.te <- read.table("UCI HAR Dataset/test/y_test.txt", col.names = "code")
X.ta <- read.table("UCI HAR Dataset/train/X_train.txt", col.names = feat$functions)
y.ta <- read.table("UCI HAR Dataset/train/y_train.txt", col.names = "code")
```

This is then used to complete the first task of the assignment, which is to merge the files:
``` {r merge}
## 1. Merging the training and test set...
X <- rbind(X.ta, X.te)
Y <- rbind(y.ta, y.te)

# Merging the subject files
sujeto <- rbind(subject.ta, subject.te)
MergeResult <- cbind(sujeto, Y, X)
```

As we create the merged result data base, then we have to look for the mean and stdev and make another data table:
``` {r merge follow-up}
## 2. Taking the mean and standard deviation for the values...
CompData <- MergeResult %>% select(subject, code, contains("mean"), contains("std"))
```

And then, we name the activities in the data set:
``` {r naming}
## 3. Naming the activities in the data set...
CompData$code <- actvs[CompData$code, 2]
```

To which we finally label the names:
``` {r labeling}
names(CompData)[2] = "activity"
names(CompData)<-gsub("Acc", "Accelerometer", names(CompData))
names(CompData)<-gsub("Gyro", "Gyroscope", names(CompData))
names(CompData)<-gsub("BodyBody", "Body", names(CompData))
names(CompData)<-gsub("Mag", "Magnitude", names(CompData))
names(CompData)<-gsub("^t", "Time", names(CompData))
names(CompData)<-gsub("^f", "Frequency", names(CompData))
names(CompData)<-gsub("tBody", "TimeBody", names(CompData))
names(CompData)<-gsub("-mean()", "Mean", names(CompData), ignore.case = TRUE)
names(CompData)<-gsub("-std()", "STD", names(CompData), ignore.case = TRUE)
names(CompData)<-gsub("-freq()", "Frequency", names(CompData), ignore.case = TRUE)
names(CompData)<-gsub("angle", "Angle", names(CompData))
names(CompData)<-gsub("gravity", "Gravity", names(CompData))
```

And we print the resulting graph!
``` {r table}
FinalData <- CompData %>%
    group_by(subject, activity) %>%
    summarise_all(funs(mean))
write.table(FinalData, "FinalData.txt", row.name=FALSE)
head(FinalData)
```

And that is it for the project!